---
title: "Palmetto Binary Logistic Regression (Task 1)"
author: "Lucas Boyd"
date: "1/28/2022"
output: 
   html_document:
     code_folding: hide
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# attach packages
library(tidyverse)
library(here)
library(broom)
library(patchwork)
library(AICcmodavg)
library(caret)
library(kableExtra)
```

## A. Overview

An overview section describing the data, the question(s) to be addressed in your analysis, and a citation of the dataset.

**Data Citation:** Data source: Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5

### Reading in and cleaning up the data
```{r}
palmetto <- read_csv(here("data", "palmetto.csv")) %>% 
  select(height, length, width, green_lvs, species) %>% # selecting for plant height (height), canopy length (length), canopy width (width), and number of green leaves (green_lvs)
# 1 = Serenoa repens, 2 = Sabal etonia
  drop_na() %>% # dropping NA values
  mutate(species_name = case_when(
    species == 1 ~ "Serenoa repens", # new column for species name
    species == 2 ~ "Sabal etonia"
  )) %>% 
  mutate(species_name = as_factor(species_name)) # changing species name to a factor for use in blr
```

# B. Visualizing differences between the two species
```{r}
# ggplots to visualize the differences in each variable between species
# effectplot could also be useful in these visualizations for blr 

p1 <-ggplot(data = palmetto, aes(x = species_name, y = green_lvs, fill = species_name)) + 
  geom_violin(scale = "count", color = "black") + # geom violin to visualize the large volume of data
geom_boxplot(color = "black", fill = NA, width = 0.1, outlier.color = NA) + # adding summary median and quartiles to figure
  stat_summary(fun=mean, # adding the mean to the figure 
               geom="point", 
               shape=20, 
               size=4, 
               color="black", 
               fill="black") +
  scale_fill_manual(values = c("orange4", "darkolivegreen")) + # customizing colors
  theme_minimal(13) + # changing theme and font size
  theme(legend.position = "none",
        axis.text.x = element_text(vjust = 5, size = 12)) + # customizing test
  labs(x = element_blank(), y = "Number of green leaves") # adding labels

# repeated above code for each variable in subsequent 3 plots

p2 <- ggplot(data = palmetto, (aes(x = species_name, y = height, fill = species_name))) +
    geom_violin(scale = "count", color = "black") +
geom_boxplot(color = "black", fill = NA, width = 0.1, outlier.color = NA) +
  stat_summary(fun=mean, 
               geom="point", 
               shape=20, 
               size=4, 
               color="black", 
               fill="black") +
  scale_fill_manual(values = c("orange4", "darkolivegreen")) +
  theme_minimal(13) +
  theme(legend.position = "none",
        axis.text.x = element_text(vjust = 5, size = 12)) +
  labs(x = element_blank(), y = "Plant Height (cm)") 

p3 <- ggplot(data = palmetto, (aes(x = species_name, y = width, fill = species_name))) +
    geom_violin(scale = "count", color = "black") +
geom_boxplot(color = "black", fill = NA, width = 0.1, outlier.color = NA) +
  stat_summary(fun=mean, 
               geom="point", 
               shape=20, 
               size=4, 
               color="black", 
               fill="black") +
  scale_fill_manual(values = c("orange4", "darkolivegreen")) +
  theme_minimal(13) +
  theme(legend.position = "none",
        axis.text.x = element_text(vjust = 5, size = 12)) +
  labs(x = element_blank(), y = "Canopy Width (cm)") 

p4 <- ggplot(data = palmetto, (aes(x = species_name, y = length, fill = species_name))) +
    geom_violin(scale = "count", color = "black") +
geom_boxplot(color = "black", fill = NA, width = 0.1, outlier.color = NA) +
  stat_summary(fun=mean, 
               geom="point", 
               shape=20, 
               size=4, 
               color="black", 
               fill="black") +
  scale_fill_manual(values = c("orange4", "darkolivegreen")) +
  theme_minimal(13) +
  theme(legend.position = "none",
        axis.text.x = element_text(vjust = 5, size = 12)) +
  labs(x = element_blank(), y = "Canopy Length (cm)") 
```

```{r}
(p1 | p2)
(p3 | p4)
```

**Figure 1** compares two palmetto species to determine feasibility of each of four variables in predicting whether a plant is *Serenoa repens* (green) or *Sabal etonia* (orange). Variables examined in the figure are: number of green leaves on the plant (top left), height of plant (cm) (top right), canopy width (cm) (bottom left) and canopy length (cm) (bottom right). Width of the shape in each figure represents the quantity of measurments at that particular value. Box endpoints indicate the 25^th^ and 75^th^ percentile values; the black line and black point within the box indicate the median and mean value for each species, respectively. Based on this visualization, it appears that number of green leaves, canopy length and canopy width will all be useful in predicting the species of the palmetto. 

**Conclusion:** Based on this visualization, it appears that number of green leaves and canopy length will be effectively in correctly classifying species, since there appears to be a noticeable difference in the mean, median and overall spread of the data. Plant height and canopy width differ only very slightly between the two species, and thus will be less useful in classifying the species of plant from the data. 

### C. Binomial logistic regression

```{r}
# storing functions of combinations of variables for use in regression
f1 <- species_name ~ height + length + width + green_lvs
f2 <- species_name ~ height + width + green_lvs

# running and storing the regressions for each formula
blr1 <- glm(formula = f1,
            data = palmetto,
            family = "binomial")

blr2 <- glm(formula = f2,
            data = palmetto,
            family = "binomial")

# taking a look at the outputs of the regression
# reference (0) = Sabal etonia
summary(blr1)
summary(blr2) 

aictab(list(blr1, blr2))
# model1 = 5194
# model2 = 5987
```

### Model Selection 

#### 10-fold cross validation

```{r}
set.seed(24) # setting seed for reproducibility

tr_ctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 10) # stroing train control, specifying 10 folds



# train the model
model1 <- train(f1, data = palmetto,
                method = "glm",
                family = "binomial",
                trControl = tr_ctrl)

model1 # accuracy = 0.917

# train the other model
model2 <- train(f2, data = palmetto,
                method = "glm",
                family = "binomial",
                trControl = tr_ctrl)

model2 # accuracy = 0.899
```

Model 1, which classifies palmetto species as a function of plant height, canopy width, canopy length and number of green leaves, performed better during 10-fold cross validation (accuracy = 91.7%, AICc = 5194) than model 2 (accuracy = 89.9%, AICc = 5987). Thus, based on the prediction accuracy results and AIC values, model 1 is preferred. 

### D. Regression results

**Table 1** shows the results of binomial logistic regression for model 1. Coefficients show the change in log odds of the resulting species being *Sabal etonia* (reference) as opposed to *Serenoa repens*.
```{r}
# storing the results of regression as a data frame
blr1_tidy <- tidy(blr1) %>%  
  mutate(p.value = case_when( # finalizing p.values for table
    p.value <0.001 ~ "<0.001"
  )) %>% 
  mutate(term = case_when( # finalizing variable names for table
         term == "(Intercept)" ~ "Intercept",
         term == "height" ~ "Height",
         term == "length" ~ "Length",
         term == "width" ~ "Width",
         term == "green_lvs" ~ "# of green leaves")) %>% 
  select(-statistic)

# finalizing table
kable(blr1_tidy, 
      col.names = c("Term", "Coefficient", "Standard Error", "P Value")) %>% 
  kable_minimal(full_width = FALSE)
  
```

### E. Model success rate

**Table 1** shows the success rate of the chosen model in classifying each plant species. 
```{r}
# storing a function for prediction accuracy
prediction_accuracy <- function(x, y) {
  accurate <- ifelse(x == y, 1, 0)
  return(accurate)
}

# creating a new dataframe with predictions based on our model
# could also use augment() here: palmetto_augment <- blr1 %>% augment(type.predict = "response") 
palmetto_success <- palmetto %>% 
  mutate(chances_sabal = predict(blr1, # adding a new column for the probability that the species is sabal, given our model
                   palmetto,
                   type = "response")) %>% 
  mutate(predicted = ifelse(chances_sabal >= .50, "Sabal etonia", "Serenoa repens")) %>%  # adding a new column that predicts the species based on the probability, using 50% as the threshold
  mutate(accurate = prediction_accuracy(species_name, predicted))

# creating a final table for model prediction accuracy for each plant

accuracy_table <- palmetto_success %>% 
  group_by(species_name) %>% # grouping by species
  summarize(total_accurate = sum(accurate), # finding total accurate predictions
            total_inaccurate = n() - sum(accurate), # total inaccurate predictions
            accuracy = (mean(accurate))*100) # accuracy %
accuracy_table
kable(accuracy_table, digits = 1, col.names = c("Palmetto Species", "Accurate", "Inaccurate", "% Accurate ")) %>% 
  kable_minimal(full_width = FALSE)


```

1-2 sentence conclusion paragraph




